---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eia2

<!-- badges: start -->
<!-- badges: end -->

The `eia2` package provides functions to download data from the [US Energy Information Administration Open Data API version 2]("https://www.eia.gov/opendata/").

Using the EIA API (via this package or other means) requires a registered API key.

The EIA provides a data browser and documentation of API concepts that is helpful to understanding usage of this package.


## Installation

You can install the development version of eia2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jameelalsalam/eia2")
```

## Authentication

Ensure your API key is available by setting it as an environment variable. You can add it to your user-level .Renviron file by:

``` r
usethis::edit_r_environ("user")
```

And adding a line such as:
EIA_KEY=<YOUR-KEY-HERE>

After restarting, running `Sys.getenv("EIA_KEY")` should return your key.


## Explore Available Data

You can browse available datasets:

```{r}
library(eia2)

eia2()
```

And iteratively explore routes within a dataset, such as:

```{r}
eia2("electricity")
```
When you get to the bottom of a route, there won't be any further routes, but you need to specify other parameters to retrieve data, such as which data columns to retrieve.

```{r}
eia2("electricity/retail-sales")
```

And finally download data such as:

```{r}
elec_retail_sales_annual_data <-
  eia2("electricity/retail-sales", frequency = "annual", data_cols = "revenue")

elec_retail_sales_annual_data
```

A full API call can also specify facet filters and different start and end dates.

```{r}
elec_retail_CO <- 
  eia2("electricity/retail-sales",
       frequency = "monthly",
       facets = list(
         stateid = c("CO", "WY")
       ),
       data_cols = c("revenue"),
       start = "2020-01",
       end = "2020-03")

elec_retail_CO
```




## Moving from API version 1

The EIA version 2 API provides a mechanism to retrieve similar data as was available via the series ID API endpoint in the version 1 API. This web form on the EIA Open Data website can help translate version 1 series IDs into version 2 routes. In addition, this package provides a function `eia1_series` which can make requests using legacy series id's.

```{r}
eia1_series("ELEC.SALES.CO-RES.A")
```



## Specifying Related Data via Parameters and Facets

The EIA version 2 API organizes dataset topics using routes, and then allows specification of various parameters and facets to narrow the specific data being requested.

...


## Acknowledgements

eia2 is based on previous work by Matthew Leonawicz on the [eia](https://github.com/ropensci/eia) package. It has benefited from many useful packages and resources including httr2, ropensci package guide, and HTTP testing for R.

